---
title: "readme"
author: "Andrew"
date: "8/18/2022"
output: rmarkdown::github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
library(tidyverse)
library(gdm)
library(raster)
library(sp)
library(rgdal)
library(colorRamps)
library(RStoolbox) # for ggRGB function

knitr::opts_chunk$set(
  fig.path = "output/readme-"
)

knitr::opts_chunk$set(
        echo = FALSE, message = FALSE, warning = FALSE
)
```

# Summary

Demonstrate a workflow / use case for passive acoustic monitoring data, with GDM

## Introduction

## Methods

Model 1:

The CSV files generated were imported into R. GDM was conducted using environmental data collected from the recording locations, to determine if compositional disimilarity was related to changes in habitat metrics. Distance to water source was calculated as a distance in meters from each study site location, based on the QLD waterways spatial layer. 

Model 2: 

A model was also conducted with raster layer spatial data obtained from the Sentinel hub platform. To determine the relationships with remotely sensed data (vegetation/greenness indices) and variation in compositional dissimilarity of bird species at the property. The indices used were SAVI and NDVI.  

Satellite images were used for the dates (what dates) when the study took place (iamges with cloud cover 0%). 

## Results

```{r}
# import the data
source("scripts/data-prep.R")
```

Species frequency table: 
```{r}
tbl_04
```

### GDM model 1
#### Detections vs vegetation survey data
```{r}
# source the gdm prep
source("scripts/gdm.R")

# fit a gdm
gdm_01 <- gdm(data=gdmTab, geo = TRUE)
summary(gdm_01)
```

```{r}
#length(gdm_01$predictors) # get ideal of number of panels
plot(gdm_01, plot.layout=c(2,3))
```

#### Detections vs remote sensing

"Environmental data can be extracted directly from rasters, assuming the x-y coordinates of sites are provided in either a site-species table (bioFormat=1) or as a x-y species list (bioFormat=2)."

```{r}
# environmental raster data for sw oz

savi <- raster("gis/sentinel/Sentinel-2-L1C_SAVI-GREEN-TO-WHITE_2017-08-29_2017-09-25_16196615.25_-3250094.53_16209828.34_-3243358.02_EPSG_3577_0_mostRecent_logo.tif", band = 2)

plot(savi)

ndvi <- raster("gis/sentinel/Sentinel-2-L1C_NDVI_2017-08-29_2017-09-25_16195224.26_-3250521.17_16211366.77_-3242695.26_EPSG_3577_0_mostRecent_logo.tif",  band = 2)

plot(ndvi)

stack_test <- stack("gis/sentinel/Sentinel-2-L1C_SAVI-GREEN-TO-WHITE_2017-08-29_2017-09-25_16196615.25_-3250094.53_16209828.34_-3243358.02_EPSG_3577_0_mostRecent_logo.tif")

plotRGB(stack_test)

newproj <- "+proj=aea +lat_0=0 +lon_0=132 +lat_1=-18 +lat_2=-36 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=61.55,-10.87,-40.19,-39.4924,-32.7221,-32.8979,-9.99400000001316 +units=m +no_defs +type=crs"

testing <- projectRaster(savi, crs = newproj)

gdmTab.rast <- formatsitepair(bioData=sppTab_Bowra, 
                              bioFormat=2, # x-y spp list
                              XColumn="X", 
                              YColumn="Y", 
                              sppColumn="common_name",
                              siteColumn="point_name",
                              predData=savi) #raster stack

gdmTab.rast.2 <- formatsitepair(bioData=sppTab_Bowra, 
                              bioFormat=2, # x-y spp list
                              XColumn="X", 
                              YColumn="Y", 
                              sppColumn="common_name",
                              siteColumn="point_name",
                              predData=ndvi) #raster stack
```

"Because some sites might not overlap with the rasters, we should check for and remove NA values from the site-pair table"

```{r}
sum(is.na(gdmTab.rast))
#gdmTab.rast <- na.omit(gdmTab.rast)
```

"Note that the formatsitepair function assumes that the coordinates of the sites are in the same coordinate system as the rasters. At present, no checking is performed to ensure this is the case. Note also that if your site coordinates are longitude-latitude that the calculation of geographic distances between sites will have errors, the size of which will depend on the geographic extent and location of your study region. We hope to deal with this in a later release, but for now you can avoid these problems by using a projected coordinate system (e.g., equidistant)."

Spatial coordinate system I am using is EPSG:3577 (X, Y cols in metadata) with reprojected raster layers from Sentinel Hub.  

Fit a GDM using the satellite data: 

```{r}
# fit a new gdm using a table with sat data only (to match rasters)

gdm.rast <- gdm(gdmTab.rast, geo=T)

summary(gdm.rast)

gdm.rast.2 <- gdm(gdmTab.rast.2, geo=T)

summary(gdm.rast.2)

transRasts <- gdm.transform(model=gdm.rast, data=savi)

raster::plot(transRasts, col=rgb.tables(1000))
```


```{r}
# Get the data from the gdm transformed rasters as a table

rastDat <- na.omit(raster::getValues(transRasts))

# The PCA can be fit on a sample of grid cells if the rasters are large

#rastDat <- raster::sampleRandom(transRasts, 50000) 

# perform the principle components analysis

pcaSamp <- prcomp(rastDat)
 
# Predict the first three principle components for every cell in the rasters

# note the use of the 'index' argument

pcaRast <- raster::predict(transRasts, pcaSamp, index=1:3)

# scale the PCA rasters to make full use of the colour spectrum

pcaRast[[1]] <- (pcaRast[[1]]-pcaRast[[1]]@data@min) /
  (pcaRast[[1]]@data@max-pcaRast[[1]]@data@min)*255
pcaRast[[2]] <- (pcaRast[[2]]-pcaRast[[2]]@data@min) /
  (pcaRast[[2]]@data@max-pcaRast[[2]]@data@min)*255
pcaRast[[3]] <- (pcaRast[[3]]-pcaRast[[3]]@data@min) /
  (pcaRast[[3]]@data@max-pcaRast[[3]]@data@min)*255

# Plot the three PCA rasters simultaneously, each representing a different colour 
#  (red, green, blue)

raster::plotRGB(pcaRast, r=1, g=2, b=3)

plot(X ~ Y, data = metadata, add = TRUE)

metadata

ggRGB(pcaRast, r=1, g=2, b=3) + geom_point(data = metadata, aes(x = X, y = Y, color = treatment))

```


```{r}

```


# Discussion

